---
title: "Brexit Market on Betfair"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
## Loading required packages
library(flexdashboard)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)

## Removing Scientific Notation
options(scipen=999)

## Loading data files
brexit.best.prices <- read.csv("brexit_best_prices.csv")
brexit.matched.volume <- read.csv("brexit_matched_volume.csv")

## Data Manipulation
brexit.best.prices$TAKEN_DATE <- ymd_hms(brexit.best.prices$TAKEN_DATE)
brexit.best.prices$SELECTION_ID <- factor(brexit.best.prices$SELECTION_ID, labels = c("Leave","Remain"))

brexit.matched.volume$TAKEN_DATE <- ymd_hms(brexit.matched.volume$TAKEN_DATE)

## Creating variables to identify last 24 hours of the market
last.date <- max(brexit.best.prices$TAKEN_DATE)
one.day.before.last.date <- last.date - 24*60*60
```

### **Market on the day of the vote** 

```{r on_day}
## Subsetting last 24 hours of the market
brexit.best.prices.last.24.hours <- brexit.best.prices %>% 
                                    filter(between(TAKEN_DATE, one.day.before.last.date, last.date))

## Plotting the last 24 hours
last.24.hours.plot <- ggplot(brexit.best.prices.last.24.hours,
                             aes(x = TAKEN_DATE,
                                 y = 1/BEST_BACK_PRICE,
                                 group = SELECTION_ID,
                                 colour = SELECTION_ID)) +
                      geom_line() +
                      theme_minimal() + 
                      xlab("Time") + 
                      ylab("Implied Probability (1/Best Price or Odds)") + 
                      scale_color_brewer(palette = "Set1", name = "Vote")

last.24.hours.plot
```

> There was incredible price fluctuation on the day of the vote (and count) where Vote Remain was tipped to be the likely outcome up until the last 6 hours. Prior to that the odds of Vote Leave begin to shorten eventually moving into favouritism. There is a tiny blip where it goes back to being outside shot, but eventually drifts into really low odds (or extremely high probability) all the way down to 1.01