---
title: "Brexit Market on Betfair"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
## Loading required packages
library(flexdashboard)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)

## Removing Scientific Notation
options(scipen=999)

## Loading data files
brexit.best.prices <- read.csv("brexit_best_prices.csv")
brexit.matched.volume <- read.csv("brexit_matched_volume.csv")

## Data Manipulation
brexit.best.prices$TAKEN_DATE <- ymd_hms(brexit.best.prices$TAKEN_DATE)
brexit.best.prices$SELECTION_ID <- factor(brexit.best.prices$SELECTION_ID, labels = c("Leave","Remain"))

brexit.matched.volume$TAKEN_DATE <- ymd_hms(brexit.matched.volume$TAKEN_DATE)

## Creating variables to identify last 24 hours of the market
last.date <- max(brexit.best.prices$TAKEN_DATE)
one.day.before.last.date <- last.date - 24*60*60
```

### **Market prior to the day of the vote**

```{r prev_months}
## Subsetting all but last 24 hours of the market
brexit.best.prices.prev.months <- brexit.best.prices %>% 
                                  filter(!between(TAKEN_DATE, one.day.before.last.date, last.date))

## Plotting the prior months of the market
prev.months.plot <- ggplot(brexit.best.prices.prev.months,
                           aes(x = TAKEN_DATE,
                               y = 1/BEST_BACK_PRICE,
                               group = SELECTION_ID,
                               colour = SELECTION_ID)) +
                    geom_line() +
                    theme_minimal() +
                    xlab("Time") + 
                    ylab("Implied Probability (1/Best Price or Odds)") +
                    scale_color_brewer(palette = "Set1", name = "Vote") 

prev.months.plot
```

> For months leading up to the last day, prior to the day of the counting, there have been a few price changes but Vote Remain has been odds on favourite for nearly a year.

### **Market on the day of the vote** 

```{r on_day}
## Subsetting last 24 hours of the market
brexit.best.prices.last.24.hours <- brexit.best.prices %>% 
                                    filter(between(TAKEN_DATE, one.day.before.last.date, last.date))

## Plotting the last 24 hours
last.24.hours.plot <- ggplot(brexit.best.prices.last.24.hours,
                             aes(x = TAKEN_DATE,
                                 y = 1/BEST_BACK_PRICE,
                                 group = SELECTION_ID,
                                 colour = SELECTION_ID)) +
                      geom_line() +
                      theme_minimal() + 
                      xlab("Time") + 
                      ylab("Implied Probability (1/Best Price or Odds)") + 
                      scale_color_brewer(palette = "Set1", name = "Vote")

last.24.hours.plot
```

> There was incredible price fluctuation on the day of the vote (and count) where Vote Remain was tipped to be the likely outcome up until the last 6 hours. Prior to that the odds of Vote Leave begin to shorten eventually moving into favouritism. There is a tiny blip where it goes back to being outside shot, but eventually drifts into really low odds (or extremely high probability) all the way down to 1.01

### **Matched Volume** 

```{r matched_volume}
## Creating Dataframe for cumulative matched volume in both GBP and AUD
brexit.matched.volume.cum.last.24.hours <- brexit.matched.volume %>% 
                                           mutate(MATCHED_GBP_CUMULATIVE = cumsum(MATCHED_GBP),
                                           MATCHED_GBP_AUD = MATCHED_GBP_CUMULATIVE * FX_RATE) %>%
                                           select(-FX_RATE, -MATCHED_GBP, - MATCHED_AUD) %>%
                                           gather("currency","value",2:3) %>%
                                           filter(between(TAKEN_DATE, one.day.before.last.date, last.date))
## Plotting Cumulative Matched Volume
matched.volume.plot <- ggplot(brexit.matched.volume.cum.last.24.hours,
                              aes(x = TAKEN_DATE,
                                  y = value,
                                  group = currency,
                                  colour = currency)) +
                       geom_line() +
                       theme_minimal() +
                       xlab("Time") +
                       ylab("Cumulative Matched Amount") +
                       scale_colour_brewer(palette = "Dark2", name = "Currency") +
                       theme(legend.position = "bottom")

matched.volume.plot
```

> Knock on effects were being felt everywhere as the Pound Sterling dropped, and one rather cheeky or funny effect was the matched volume amount shown on Betfair Australia's website - in AUD, was going lower even as more money was coming in and getting matched. 