---
title: "Sports on Betfair"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: flatly
---

```{r setup, include=FALSE}
## Loading required packages
library(flexdashboard)
library(leaflet)
library(plotly)
library(dplyr)
library(tidyr)
library(ggplot2)
library(dygraphs)
library(lubridate)
library(xts)

## Removing scientific notation
options(scipen = 999)

## Loading data files 
sports.geo <- read.csv("sports_geo.csv")
sports.ts <- read.csv("sports_ts.csv")
sports.in.play <- read.csv("sports_in_play.csv")

## Data Wrangling

## Converting to sports.ts data frame to a xts time series object after some data wrangling
sports.ts <- sports.ts %>%
             mutate(calendar_dt = ymd_hms(calendar_dt)) %>%
             spread(sport_name, turnover_gbp) 

rownames(sports.ts) <- sports.ts$calendar_dt
sports.ts[is.na(sports.ts)] <- 0

sports.ts <- sports.ts %>% 
             select(-calendar_dt) %>%
             as.matrix() %>%
             as.xts(dateFormat = 'Date')

```

Aussie Rules
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### **Seasonality**

```{r afl_ts}
## Using dygraphs to plot time series of Aussie Rules turnover since 2013
dygraph(sports.ts[,"Australian Rules"]) %>%
  dyOptions(maxNumberWidth = 20) %>%
  dyAxis("y", label = "Turnover")
```

Row {data-height=600}
-----------------------------------------------------------------------

### **Customers Location**

```{r aussie_rules_geo}
## Using Leaflet Widget to plot map of Aussie Rules customers based on IP Address
leaflet(sports.geo %>%
        filter(sport_name == "Australian Rules")) %>%
  addTiles() %>%
  addCircleMarkers(lat = ~latitude,
                   lng = ~longitude, 
                   clusterOptions = markerClusterOptions())
```


### **Pre-play vs In-Play**

```{r aussie_rules_in_play}
## Using plotly to create scatter plot showing relationship between turnover generated 
## pre play (before game starts) and turnover generated in play (after game starts) in Aussie Rules
plot_ly(sports.in.play %>%
          filter(sport_name == "Australian Rules"),
        x = pre_play_turnover,
        y = in_play_turnover,
        mode = "markers",
        type = "scatter") %>%
  layout(xaxis = list(
           title = "Pre Play Turnover"
         ),
         yaxis = list(
           title = "In Play Turnover"
         )
  )
```


Rugby League
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### **Seasonality**

```{r rugby_league_ts}
## Using dygraphs to plot time series of Rugby League turnover since 2013
dygraph(sports.ts[,"Rugby League"]) %>%
  dyOptions(maxNumberWidth = 20) %>%
  dyAxis("y", label = "Turnover")
```

Row {data-height=600}
-----------------------------------------------------------------------

### **Customers Location**

```{r rugby_league_geo}
## Using Leaflet Widget to plot map of Rugby League customers based on IP Address
leaflet(sports.geo %>%
        filter(sport_name == "Rugby League")) %>%
  addTiles() %>%
  addCircleMarkers(lat = ~latitude,
                   lng = ~longitude, 
                   clusterOptions = markerClusterOptions())
```

### **Pre-play vs In-Play**

```{r rugby_league_in_play}
## Using plotly to create scatter plot showing relationship between turnover generated 
## pre play (before game starts) and turnover generated in play (after game starts) in Rugby League
plot_ly(sports.in.play %>%
          filter(sport_name == "Rugby League"),
        x = pre_play_turnover,
        y = in_play_turnover,
        mode = "markers",
        type = "scatter") %>%
  layout(xaxis = list(
           title = "Pre Play Turnover"
         ),
         yaxis = list(
           title = "In Play Turnover"
         )
  )
```

Rugby Union
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### **Seasonality**

```{r rugby_union_ts}
## Using dygraphs to plot time series of Rugby Union turnover since 2013
dygraph(sports.ts[,"Rugby Union"]) %>%
  dyOptions(maxNumberWidth = 20) %>%
  dyAxis("y", label = "Turnover")
```

Row {data-height=600}
-----------------------------------------------------------------------

### **Customers Location**

```{r rugby_union_geo}
## Using Leaflet Widget to plot map of Rugby Union customers based on IP Address
leaflet(sports.geo %>%
        filter(sport_name == "Rugby Union")) %>%
  addTiles() %>%
  addCircleMarkers(lat = ~latitude,
                   lng = ~longitude, 
                   clusterOptions = markerClusterOptions())
```

### **Pre-play vs In-Play**

```{r rugby_union_in_play}
## Using plotly to create scatter plot showing relationship between turnover generated 
## pre play (before game starts) and turnover generated in play (after game starts) in Rugby Union
plot_ly(sports.in.play %>%
          filter(sport_name == "Rugby Union"),
        x = pre_play_turnover,
        y = in_play_turnover,
        mode = "markers",
        type = "scatter") %>%
  layout(xaxis = list(
           title = "Pre Play Turnover"
         ),
         yaxis = list(
           title = "In Play Turnover"
         )
  )
```

Cricket
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### **Seasonality**

```{r cricket_ts}
## Using dygraphs to plot time series of Cricket turnover since 2013
dygraph(sports.ts[,"Cricket"]) %>%
  dyOptions(maxNumberWidth = 20) %>%
  dyAxis("y", label = "Turnover")
```

Row {data-height=600}
-----------------------------------------------------------------------

### **Customers Location**

```{r cricket_geo}
## Using Leaflet Widget to plot map of Cricket customers based on IP Address
leaflet(sports.geo %>%
        filter(sport_name == "Cricket")) %>%
  addTiles() %>%
  addCircleMarkers(lat = ~latitude,
                   lng = ~longitude, 
                   clusterOptions = markerClusterOptions())
```

### **Pre-play vs In-Play**

```{r cricket_in_play}
## Using plotly to create scatter plot showing relationship between turnover generated 
## pre play (before game starts) and turnover generated in play (after game starts) in Cricket
plot_ly(sports.in.play %>%
          filter(sport_name == "Cricket"),
        x = pre_play_turnover,
        y = in_play_turnover,
        mode = "markers",
        type = "scatter") %>%
  layout(xaxis = list(
           title = "Pre Play Turnover"
         ),
         yaxis = list(
           title = "In Play Turnover"
         )
  )
```

Tennis
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### **Seasonality**

```{r tennis_ts}
## Using dygraphs to plot time series of Tennis turnover since 2013
dygraph(sports.ts[,"Tennis"]) %>%
  dyOptions(maxNumberWidth = 20) %>%
  dyAxis("y", label = "Turnover")
```

Row {data-height=600}
-----------------------------------------------------------------------

### **Customers Location**

```{r tennis_geo}
## Using Leaflet Widget to plot map of Tennis customers based on IP Address
leaflet(sports.geo %>%
        filter(sport_name == "Tennis")) %>%
  addTiles() %>%
  addCircleMarkers(lat = ~latitude,
                   lng = ~longitude, 
                   clusterOptions = markerClusterOptions())
```

### **Pre-play vs In-Play**

```{r tennis_in_play}
## Using plotly to create scatter plot showing relationship between turnover generated 
## pre play (before game starts) and turnover generated in play (after game starts) in Tennis
plot_ly(sports.in.play %>%
          filter(sport_name == "Tennis"),
        x = pre_play_turnover,
        y = in_play_turnover,
        mode = "markers",
        type = "scatter") %>%
  layout(xaxis = list(
           title = "Pre Play Turnover"
         ),
         yaxis = list(
           title = "In Play Turnover"
         )
  )
```

Soccer
=======================================================================

Row {data-height=400}
-----------------------------------------------------------------------

### **Seasonality**

```{r soccer_ts}
## Using dygraphs to plot time series of Soccer turnover since 2013
dygraph(sports.ts[,"Soccer"]) %>%
  dyOptions(maxNumberWidth = 20) %>%
  dyAxis("y", label = "Turnover")
```

Row {data-height=600}
-----------------------------------------------------------------------

### **Customers Location**

```{r soccer_geo}
## Using Leaflet Widget to plot map of Soccer customers based on IP Address
leaflet(sports.geo %>%
        filter(sport_name == "Soccer")) %>%
  addTiles() %>%
  addCircleMarkers(lat = ~latitude,
                   lng = ~longitude, 
                   clusterOptions = markerClusterOptions())
```

### **Pre-play vs In-Play**

```{r soccer_in_play}
## Using plotly to create scatter plot showing relationship between turnover generated 
## pre play (before game starts) and turnover generated in play (after game starts) in Soccer
plot_ly(sports.in.play %>%
          filter(sport_name == "Soccer"),
        x = pre_play_turnover,
        y = in_play_turnover,
        mode = "markers",
        type = "scatter") %>%
  layout(xaxis = list(
           title = "Pre Play Turnover"
         ),
         yaxis = list(
           title = "In Play Turnover"
         )
  )
```